---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
#---- Packages ----
library(readxl)
library(ggplot2)
library(astrochron)
library(viridis)
library(WaverideR)
library(cowplot)
```

# Data Preparation

```{r}
df <- read.csv("/Users/silkemaus/Downloads/PhD/April26/cyclo_EGU/1408_1410.csv")
#colnames(df)

series <- data.frame(
  depth = df$Depth.comp..mcd...two.site.composite..Depth.CCS....,
  TiAl = df$Ti_Al,
  Age = df$Age_Ma
)
colnames(series) <- c("depth", "TiAl", "Age")

series <- subset(series, depth >= 50 & depth <= 240)

range(series$Age)
```

Depth Span: 190 m Age Span: 6.39796 Myr --\> 6397.96 kyr

--\> SR = 190m/6.39796 Myr = 29.6970 m/Myr --\> 2.97 cm/kyr

# Interpolation, detrending

```{r}
# remove NA
series <- na.omit(series)

# detrend
loess_fit <- loess(TiAl ~ depth, data = series, span = 0.2)
series$TiAl_dt <- series$TiAl - predict(loess_fit)

# standardize
series$TiAl_std <- scale(series$TiAl_dt)

# prepare clean data
series_clean <- data.frame(
  depth = series$depth,
  TiAl = as.numeric(series$TiAl_std)
)

# remove outliers HERE (before interpolation)
q <- quantile(series_clean$TiAl, c(0.01, 0.99))
series_clean <- series_clean[
  series_clean$TiAl > q[1] & series_clean$TiAl < q[2],
]

# NOW interpolate (this restores uniform spacing)
series_int <- linterp(series_clean, dt = 0.02)
```

```{r}
plot(series$depth, series$TiAl_std,
     type = "l",
     xlab = "Depth (mcd)",
     ylab = "Standardized Ti/Al",
     main = "Detrended & Standardized Ti/Al")
grid()
```

# MTM Spectral Analysis

-   1/405 → long eccentricity

-   1/130 → eccentricity

-   1/33 → obliquity

-   1/20 → precession

ASM converts cycles/m ↔ cycles/kyr using sedimentation rate

periodogram(series_int)

series_det \<- detrend(series_int, method="loess", span=0.2)

```{r}

TiAl_mtm <- mtm(series_int, tbw=2, ntap=3, output =1)
names(TiAl_mtm)
```

Rednoise_CL ≳ 95% → robust signal very strong periods: ID / Frequency /
Period / Harmonic_CL / Rednoise_CL 1 0.07051741 14.1809 90.69917 100
--\> 473 kyr 2 0.08314739 12.02684 99.98571 99.99799 --\> 400 kyr 3
0.09577738 10.44088 98.52804 100 --\> 347 kyr 4 0.1073549 9.314902
98.40109 95.10091\
5 0.1315623 7.60096 92.11312 99.92148 --\> 253 kyr 6 0.1631373 6.129806
94.17595 98.34598

## 8 0.1905023 5.249282 95.38511 98.86775 --\> 175 kyr

below period is too small for orbital forcing

do they correspond to eccentricity, obliquity, precession?

use SR from earlier: SR = 2.97 cm/kyr = 0.0297 m/kyr

eg: 14.2/0.0297 = 473 kyr (could be long ecc) no obliquity, precession

--\> refine and estimate SR more precisely SR=12/405≈0.0296m/kyr

#bandpass filter for main band: 5-15m 15m --\> 0.067 (f=1/lambda) 5m
--\> 0.20

```{r}

bp <- bandpass(
  series_int,
  flow = 0.067,
  fhigh = 0.20,
  win = 1,
  output = 1
)
```

```{r}
plot(bp[,1], bp[,2],
     type="l",
     xlab="Depth (mcd)",
     ylab="Filtered signal",
     main="Bandpass: 5–15 m (eccentricity band)")
grid()
```

# Eha

This checks if your signal is: stable through depth (true orbital) or
random patches (noise)

```{r}
eha_res <- eha(series_int, win=20, step=1)

# remove problematic values
eha_res$z[!is.finite(eha_res$z)] <- NA
```

```{r}
summary(series_int)
head(series_int)
```

```{r}
periodogram(series_int)
```

Plot Spectrum

```{r}
plot(TiAl_mtm$Frequency,
     TiAl_mtm$Power,
     type = "l",
     xlab = "Frequency (cycles/m)",
     ylab = "Power")

```

ASM braucht: \# 1. Frequenzen aus MTM (cycles/m) freq_TiAl \<-
TiAl_mtm\$Frequency \# 2. Rayleigh frequency (aus Periodogramm) Rayleigh
frequency: 0.005262604 \# 3. Nyquist frequency (aus Periodogramm)
Nyquist frequency: 25 \# 4. Orbital target frequencies (cycles/kyr)
mittleres Eozän target_pre \<- c(1/23, 1/21, 1/19) target_ob \<- c(1/41)
target_ecc \<- c(1/125, 1/100, 1/95) target_405 \<- c(1/405) \# 5.
Suchbereich für Sedimentationsrate

relevante peak frequenzen von mtm in cycles/m --\> transform to
wavelength lambda = 1/f freq_sig \<- c(0.07051741, 0.08314739,
0.09577738, 0.1073549, 0.1315623, 0.1631373) lambda_sig \<- c(14.2,
12.0, 10.4, 9.31, 7.6, 6.1)

```{r}
#Do these correspond to orbital cycles?
#To test that, estimate sedimentation rate (SR): SR = cycle thickness (m)/orbital period (kyr)
# eg. Precession: 14.2 m / 21 kyr ≈ 0.68 m/kyr (all around ~0.5–0.7 m/kyr)
# eg. eccentricity: 14.2 / 100 = 0.14 m/kyr (differ too much --> not plausible)
```

3 precession peaks\_ 10.4 m / 19 ≈ 0.55 12.0 m / 21 ≈ 0.57 14.2 m / 23 ≈
0.62 give good SR

```{r}
freq_TiAl <- TiAl_mtm$Frequency
ray <- 0.005262604
nyq <- 25
target_pre <- c(1/23, 1/21, 1/19)
target_ob <- c(1/41)
target_ecc <- c(1/125, 1/100, 1/95)
target_405 <- c(1/405)

# relevante peak frequenzen in cycles/m
freq_sig <- c(0.0705, 0.0831, 0.0958)
```

Precession

```{r}
TiAl_asm_pre <- asm(freq = freq_sig,
                    target = target_pre,
                    rayleigh = ray,
                    nyquist = nyq,
                    fper = rep(0.3, length(target_pre)),
                    numsed = 100,
                    sedmin = 1,    
                    sedmax = 50,  
                    iter = 500,
                    genplot = TRUE,
                    output = TRUE,
                    linLog = 1)
```

```{r}
SR_cm <- 51.79602
SR_m  <- SR_cm / 100   # = 0.52 m/kyr

# depth --> time

depth <- series_int[,1]
signal <- series_int[,2]

time_kyr <- depth / SR_m

# new df
series_time <- cbind(time_kyr, signal)

dt_time <- 0.02 / SR_m
dt_time

TiAl_mtm_time <- mtm(series_time,
                     tbw = 2,
                     ntap = 3,
                     output = 1)
```

```{r}
head(series_time)
dt_time


freq_time <- TiAl_mtm_time$Frequency
power_time <- TiAl_mtm_time$Power

valid <- freq_time > 0

period_kyr <- 1 / freq_time[valid]
power_time <- power_time[valid]

range(period_kyr)

plot(period_kyr,
     power_time,
     type = "l",
     xlim = c(0, 50),
     xlab = "Period (kyr)",
     ylab = "Power")
```

```{r}

```
