---
title: "R Notebook"
output: html_notebook
---

```{r}
#setwd("/Users/silkemaus/Downloads/PhD/cyclo_Al:Ti/testdaten")
#getwd()

#---- Packages ----
library(readxl)
library(ggplot2)
library(astrochron)
library(viridis)
library(WaverideR)
library(cowplot)
```

# Data Preparation

```{r}
df <- read.csv("/Users/silkemaus/Downloads/PhD/Germany/astrochron/XRF_U1410_with_AlTi.csv")
#colnames(df)
series <- data.frame(
  depth = df$Depth.comp..mcd...Cappelli.et.al..2019..CCSF..,
  TiAl = df$Ti_Al,
  Age = df$Age..Ma
)
colnames(series) <- c("depth", "TiAl", "Age")

series <- subset(series, depth >= 200 & depth <= 250)

#range(series$Age)
```
Depth Span: 50m
Age Span: 456 kyr

--> SR = 50m/0.456Ma = 109.6 m/Ma --> 10.96 cm/kyr

```{r}
plot(series$depth, series$TiAl,
     type = "l",
     xlab = "Depth (mcd)",
     ylab = "Ti/Al",
     main = "U1410 Ti/Al vs Depth")
grid()
```

# Interpolation

```{r}
# interpolate
series_int <- linterp(series)
```

# MTM Spectral Analysis

-   1/405 → long eccentricity

-   1/130 → eccentricity

-   1/33 → obliquity

-   1/20 → precession

ASM converts cycles/m ↔ cycles/kyr using sedimentation rate

periodogram(series_int)

series_det \<- detrend(series_int, method="loess", span=0.2)

```{r}

TiAl_mtm <- mtm(series_int, tbw=2, ntap=3, output =1)
names(TiAl_mtm)
```

0.5430712   1.841379   95.1064   97.00803 

Period  1.841379 m
Sedimentationrate from earlier: 0.1096 m/kyr

gives: 16.8 kyr

--> close to precession (19 - 23 kyr)



```{r}
periodogram(series_int)
```

Plot Spectrum

```{r}
plot(TiAl_mtm$Frequency,
     TiAl_mtm$Power,
     type = "l",
     xlab = "Frequency (cycles/m)",
     ylab = "Power")

```
ASM braucht:
# 1. Frequenzen aus MTM (cycles/m)
freq_TiAl <- TiAl_mtm$Frequency
# 2. Rayleigh frequency (aus Periodogramm)
Rayleigh frequency: 0.04681648 
# 3. Nyquist frequency (aus Periodogramm)
Nyquist frequency: 25 
# 4. Orbital target frequencies (cycles/kyr)
mittleres Eozän
target_pre <- c(1/23, 1/21, 1/19)
target_ob <- c(1/41)
target_ecc <- c(1/125, 1/100, 1/95)
target_405 <- c(1/405)
# 5. Suchbereich für Sedimentationsrate

```{r}
freq_TiAl <- TiAl_mtm$Frequency
ray <- 0.04681648
nyq <- 25
target_pre <- c(1/23, 1/21, 1/19)
target_ob <- c(1/41)
target_ecc <- c(1/125, 1/100, 1/95)
target_405 <- c(1/405)

# relevante peak frequenzen
freq_sig <- c(0.5430712, 1.179775, 14.17603 , 14.897, 15.88015, 16.25468 , 18.33333)
```


Precession

```{r}
TiAl_asm_pre <- asm(freq = freq_sig,
                    target = target_pre,
                    rayleigh = ray,
                    nyquist = nyq,
                    fper = rep(0.3, length(target_pre)),
                    numsed = 100,
                    sedmin = 5,
                    sedmax = 15,
                    iter = 500,
                    genplot = TRUE,
                    output = TRUE,
                    linLog = 1)
```

```{r}
SR_cm <- 9.3
SR_m  <- SR_cm / 100   # = 0.093 m/kyr

# depth --> time

depth <- series_int[,1]
signal <- series_int[,2]

time_kyr <- depth / SR_m

# new df
series_time <- cbind(time_kyr, signal)

dt_time <- 0.02 / SR_m
dt_time

TiAl_mtm_time <- mtm(series_time,
                     tbw = 2,
                     ntap = 3,
                     output = 1)
```
```{r}
head(series_time)
dt_time
range(period_kyr)

freq_time <- TiAl_mtm_time$Frequency
power_time <- TiAl_mtm_time$Power

valid <- freq_time > 0

period_kyr <- 1 / freq_time[valid]
power_time <- power_time[valid]

plot(period_kyr,
     power_time,
     type = "l",
     xlim = c(0, 50),
     xlab = "Period (kyr)",
     ylab = "Power")
```
```{r}

```

